### 2019-02-20 post-linearity-2018/LAM.tex
Working on the proof oflemma 29.
Removed SN, because we lack a citation that proves SN for LNL.

Similar as the sequent calculus, the term assignment for CNC logic is also
composed of two logics; intuitionistic linear logic on the left, denoted by
$\cat{C}$, and the Lambek calculus on the right, denoted by $\cat{L}$. The
two systems by a pair of monoidal adjoint functors
$\cat{C}:\func{F} \dashv \func{G}:\cat{L}$. The syntax for types and
contexts we use in the term assignment is the same as in the sequent
calculus. The rest of the syntax for the term assignment is defined as
follows.
\begin{definition}
  \label{def:Lambek-syntax}
  The following grammar describes the syntax of the term assignment of the
  CNC logic:
  \begin{center}\vspace{-3px}\small
    \begin{math}
      \begin{array}{lll}        
        \text{($\cat{C}$-Terms)} & [[t]] ::= [[x]] \mid [[trivT]] \mid [[t1 (*) t2]] \mid [[let t1 : X be q in t2]] \mid [[\x:X.t]] \mid [[t1 t2]] \mid [[ex t1 , t2 with x1 , x2 in t3]] \mid [[Gf s]]\\
        \text{($\cat{L}$-Terms)} & [[s]] ::= [[x]] \mid [[trivS]] \mid [[s1 (>) s2]] \mid [[let s1 : A be p in s2]] \mid [[let t : X be q in s]] \mid [[\l x : A.s]] \mid [[\r x : A . s]] \\
        & \,\,\,\,\,\,\,\,\,\mid [[appl s1 s2]] \mid [[appr s1 s2]] \mid [[F t]]\\        
        \text{($\cat{C}$-Patterns)} & [[q]] ::= [[trivTp]] \mid [[x]] \mid [[q1 (*) q2]] \mid [[Gf p]]\\
        \text{($\cat{L}$-Patterns)} & [[p]] ::= [[trivSp]] \mid [[x]] \mid [[p1 (>) p2]] \mid [[F q]]\\        
        \text{($\cat{C}$-Typing Judgment)} & [[I |-c t : X]]\\
        \text{($\cat{L}$-Typing Judgment)} & [[G |-l s : A]]\\
      \end{array}
    \end{math}
  \end{center}
\end{definition}

@@

We also proved that the sequent calculus formalization given in
Figure~\ref{fig:CNC-sequent-calculus} is equivalent to the typing rules (or
else called the natural deduction formalization) given in
Figure~\ref{fig:CNC-typing-rules} are equivalent, as stated in the following
theorem.
\begin{theorem}
  \label{thm:sc-nd-equiv}
  The sequent calculus ($\mathit{SC}$) and natural deduction ($\mathit{ND}$)
  formalizations for CNC logic are equivalent in the sense that there are
  two mappins $N:\mathit{SC}\rightarrow\mathit{ND}$ and
  $S:\mathit{ND}\rightarrow\mathit{SC}$ that map each rule in $\mathit{SC}$
  to a proof in $\mathit{ND}$, and each rule in $\mathit{ND}$ to a proof
  in $\mathit{SC}$, respectively.
\end{theorem}
\begin{proof}
  The proof is done case by case on each rule in the sequence calculus and
  natural deduction formalizations. It is obvious that the axioms in one
  formalization can be mapped to the axioms in the other. The introduction
  rules in $\mathit{ND}$ are mapped to the right rules in $\mathit{SC}$, and
  vice versa. The elimination rules and lefts rules are mapped to each other
  with some fiddling. For instance, the elimination rule for the
  non-commutative tensor is mapped to the following proof in $\mathit{SC}$:
  \begin{center}
    \scriptsize
    \begin{math}
      $$\mprset{flushleft}
      \inferrule* [right={\scriptsize $\ElledruleTXXcutName$}] {
        {[[I |-c t1 : X (*) Y]]} \\
        $$\mprset{flushleft}
        \inferrule* [right={\scriptsize $\ElledruleTXXtenLName$}] {
          {[[P1, x : X, y : Y, P2 |-c t2 : Z]]}
        }{[[P1, z : X (*) Y, P2 |-c let z : X (*) Y be x (*) y in t2 : Z]]}
      }{[[P1, I, P2 |-c [t1 / z](let z : X (*) Y be x (*) y in t2) : Z]]}
    \end{math}
  \end{center}
  The full proof is in Appendix~\ref{app:sc-nd-equiv}.
\end{proof}

Finally, we proved strong normalization of CNC logic via a translation to
LNL logic. The proof is in Appendix~\ref{app:strong-normalization}. We first
define a function $CL$ that maps our term assignment for natural deduction
to Benton's term assignment, and then shows that if there is reduction
$[[t1 ~> t2]]$ (resp. $[[s1 ~> s2]]$) in CNC, then there is a reduction
$CL(t_1)\leadsto CL(t_2)$ (resp. $CL(s_1)\leadsto CL(s_2)$) in LNL. Recall
that a LNL model consists of an adjunction $F:\cat{C}\dashv\cat{L}:G$ in
which $\cat{C}$ is a cartesian closed category and $\cat{L}$ is a SMCC.
The basic idea of mapping a LAM model to a LNL model is mapping the SMCC in
LAM to the cartesian closed category in LNL, and the Lambek category in LAM
to the SMCC in LNL. For instance, $s_1\triangleright s_2$ in CNC, where
$[[s1]]$ and $[[s2]]$ are atomic terms, maps to $s_1\otimes s_2$ in LNL.

>>>>

Similar as the sequent calculus, the term assignment for CNC logic is
also composed of two logics; intuitionistic linear logic on the left,
denoted by $\cat{C}$, and the Lambek calculus on the right, denoted by
$\cat{L}$. The syntax for types and contexts we use in the term
assignment is the same as in the sequent calculus. The rest of the
syntax for the term assignment is defined as follows.
\begin{definition}
  \label{def:Lambek-syntax}
  The following grammar describes the syntax of the term assignment of the
  CNC logic:
  \begin{center}\vspace{-3px}\small
    \begin{math}
      \begin{array}{lll}        
        \text{($\cat{C}$-Terms)} & [[t]] ::= [[x]] \mid [[trivT]] \mid [[t1 (*) t2]] \mid [[let t1 : X be q in t2]] \mid [[\x:X.t]] \mid [[t1 t2]] \mid [[ex t1 , t2 with x1 , x2 in t3]] \mid [[Gf s]]\\
        \text{($\cat{L}$-Terms)} & [[s]] ::= [[x]] \mid [[trivS]] \mid [[s1 (>) s2]] \mid [[let s1 : A be p in s2]] \mid [[let t : X be q in s]] \mid [[\l x : A.s]] \mid [[\r x : A . s]] \\
        & \,\,\,\,\,\,\,\,\,\mid [[appl s1 s2]] \mid [[appr s1 s2]] \mid [[F t]]\\        
        \text{($\cat{C}$-Patterns)} & [[q]] ::= [[trivTp]] \mid [[x]] \mid [[q1 (*) q2]] \mid [[Gf p]]\\
        \text{($\cat{L}$-Patterns)} & [[p]] ::= [[trivSp]] \mid [[x]] \mid [[p1 (>) p2]] \mid [[F q]]\\        
        \text{($\cat{C}$-Typing Judgment)} & [[I |-c t : X]]\\
        \text{($\cat{L}$-Typing Judgment)} & [[G |-l s : A]]\\
      \end{array}
    \end{math}
  \end{center}
\end{definition}

@@

We also proved that the sequent calculus formalization given in
Figure~\ref{fig:CNC-sequent-calculus} is equivalent to the typing rules (or
else called the natural deduction formalization) given in
Figure~\ref{fig:CNC-typing-rules} are equivalent, as stated in the following
theorem.
\begin{theorem}
  \label{thm:sc-nd-equiv}
  The sequent calculus ($\mathit{SC}$) and natural deduction ($\mathit{ND}$)
  formalizations for CNC logic are equivalent in the sense that there are
  two mappings $N:\mathit{SC}\rightarrow\mathit{ND}$ and
  $S:\mathit{ND}\rightarrow\mathit{SC}$ that map each rule in $\mathit{SC}$
  to a proof in $\mathit{ND}$, and each rule in $\mathit{ND}$ to a proof
  in $\mathit{SC}$, respectively.
\end{theorem}
\begin{proof}
  The proof is done case by case on each rule in the sequence calculus and
  natural deduction formalizations. It is obvious that the axioms in one
  formalization can be mapped to the axioms in the other. The introduction
  rules in $\mathit{ND}$ are mapped to the right rules in $\mathit{SC}$, and
  vice versa. The elimination rules and lefts rules are mapped to each other
  with some fiddling. For instance, the elimination rule for the
  non-commutative tensor is mapped to the following proof in $\mathit{SC}$:
  \begin{center}
    \scriptsize
    \begin{math}
      $$\mprset{flushleft}
      \inferrule* [right={\scriptsize $\ElledruleTXXcutName$}] {
        {[[I |-c t1 : X (*) Y]]} \\
        $$\mprset{flushleft}
        \inferrule* [right={\scriptsize $\ElledruleTXXtenLName$}] {
          {[[P1, x : X, y : Y, P2 |-c t2 : Z]]}
        }{[[P1, z : X (*) Y, P2 |-c let z : X (*) Y be x (*) y in t2 : Z]]}
      }{[[P1, I, P2 |-c [t1 / z](let z : X (*) Y be x (*) y in t2) : Z]]}
    \end{math}
  \end{center}
  The full proof is in Appendix~\ref{app:sc-nd-equiv}.
\end{proof}

%% \noindent
%% Finally, we have strong normalization of CNC logic.
%% 
%% \begin{theorem}[Strong Normalization]
%%   \label{thm:strong-normalization}
%%   CNC logic is strongly normalizing.
%% \end{theorem}
%% \begin{proof}
%%   The proof is via a translation to LNL logic, which has been proved to be
%%   strongly normalizing \cite{Mellies:2009}. The proof is in
%%   Appendix~\ref{app:strong-normalization}. We first define a function
%%   $CL$ that maps our term assignment for natural deduction to Benton's
%%   term assignment, and then shows that if there is reduction $[[t1 ~>  t2]]$ (resp. $[[s1 ~> s2]]$) in CNC, then there is a reduction
%%   $CL(t_1)\leadsto CL(t_2)$ (resp. $CL(s_1)\leadsto CL(s_2)$) in
%%   LNL. Recall that a LNL model consists of an adjunction
%%   $F:\cat{C}\dashv\cat{L}:G$ in which $\cat{C}$ is a Cartesian closed
%%   category and $\cat{L}$ is a SMCC.  The basic idea of mapping a LAM
%%   model to a LNL model is mapping the SMCC in LAM to the Cartesian
%%   closed category in LNL, and the Lambek category in LAM to the SMCC in
%%   LNL. For instance, $s_1\triangleright s_2$ in CNC, where $[[s1]]$ and
%%   $[[s2]]$ are atomic terms, maps to $CL(s_1)\otimes CL(s_2)$ in LNL.
%% \end{proof}
